---
title: "DemsWorldMap"
output: html_document
encoding: 'utf-8'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Load dataset, preprocess, keep only relevant columns
``` {r LoadData}
library(xlsx)

Sys.setlocale(category="LC_ALL", locale = "en_US.UTF-8")

#Load the dataset 
df <- read.xlsx("DemsLangDataset.xlsx", sheetIndex=1, header = T)

# Keep only relevant columns / rows
df <- df[,c('LANGUAGE', 'GEODRYER', 'GEONICHOLS', 'STOCK', 'COORDINATES', 'DISTANCE.CONTRAST', 'ORIENTATION')]
df <- df[complete.cases(df), ]

head(df)
```

### Split coordinates into Latitude and Longitude and encode them as negative/positive coordinates
``` {r SplitCoordinates}
library(tidyr)

# Split coordinates into latitude and longitude
vars = c('Lat', 'Long')
df <- separate(df, COORDINATES, into = vars, sep = " ")

# Create columns with index for positive/negative longitude and latitude based on South/North and East/West indices 
df$DirLat <- ifelse(grepl('S', df$Lat) == TRUE, -1, 1)
df$DirLong <- ifelse(grepl('E', df$Long) == TRUE, 1, -1)

# Get rid of cardinal points (info stored in DirLat and DirLong)
df$Lat <- gsub("['S]", "", df$Lat)
df$Lat <- gsub("['N]", "", df$Lat)
df$Long <- gsub("['E]", "", df$Long)
df$Long <- gsub("['W]", "", df$Long)

head(df)
```

### Split both Latitude and Longitude into Degrees and Minutes
```{r SplitGradesPrimes}

# Get rid of some character noise
df$Long <- gsub(",", "", df$Long)
df$Lat <- gsub(",", "", df$Lat)

# Create columns for split
sub_Lat <- c('Lat_Deg', 'Lat_Min')
sub_Long <- c('Long_Deg', 'Long_Min')

# Split degrees and minutes based on comma
df <- separate(df, Lat, into = sub_Lat, sep = "°")
df <- separate(df, Long, into = sub_Long, sep ="°")

# Some minutes rows have no values because their coordinate is an integer. Just set them to 0.
df$Lat_Min[df$Lat_Min == ''] <- 0
df$Long_Min[df$Long_Min == ''] <- 0

head(df)
```

## Now take it all back to two Latitude and Longitude column in degrees + decimals
``` {r CoordToDecimals}

# Convert minutes to decimals
df$Lat_Min <- as.numeric(df$Lat_Min) / 60
df$Long_Min <- as.numeric(df$Long_Min) / 60

# Reformat stuff 
df$Lat_Deg <- as.numeric(df$Lat_Deg)
df$Long_Deg <- as.numeric(df$Long_Deg)

# South / North to pos neg 
df$Lat_Deg <- df$Lat_Deg * df$DirLat
df$Lat_Min <- df$Lat_Min * df$DirLat
df$Long_Deg <- df$Long_Deg * df$DirLong
df$Long_Min <- df$Long_Min * df$DirLong

# Sum degrees and decimal minutes
df$Lat <- df$Lat_Deg + df$Lat_Min
df$Long <- df$Long_Deg + df$Long_Min

head(df)
```

### Some more processing + split into categories based on contrast number
``` {r}

library(plyr)

# Get rid of all useless columns and rename
df <- df[, c("LANGUAGE",'GEODRYER', 'GEONICHOLS', 'STOCK', "Lat", "Long", "DISTANCE.CONTRAST", 'ORIENTATION')]
colnames(df) <- c("Lang", 'GeoDryer', 'GeoNichols', 'Stock', "Lat", "Long", "Contrasts", "Orientation")

# Apparently it has to be done in this cumbersome way, yes
df$Contrasts <- as.character(df$Contrasts)
df <- subset(df, df$Contrasts %in% c('1','2','3','4','5','6'))
df$Contrasts <- as.numeric(df$Contrasts)

# Split into three categories (less, equal or more than 3 contrast)
df$Categories[df$Contrasts < 3] <- -1
df$Categories[df$Contrasts == 3] <- 0
df$Categories[df$Contrasts > 3] <- 1

# Categories as factor (for the map thing)
df$Categories <- as.factor(df$Categories)

# Rename
df$Categories <- revalue(df$Categories, c("-1" ="Less than 3", "0" = 'Exactly 3', "1" = 'More than 3'))

head(df)
```

### DRAW THE MAP!
``` {r WorldMap}

library(ggmap)
library(maps)
library(maptools)

#df$Contrasts <- as.factor(df$Contrasts)

# Create map dummy object
world_map <- NULL

# Set type of map, colours of boundaires and fill
mapWorld <- borders("world", colour="gray50", fill="gray50")

# Store plot into world_map
world_map <- ggplot() +   mapWorld

# Set colorPalette and take n colors from it
# colfunc<-colorRampPalette(c("green","yellow","red"))
#col = colfunc(3)

# Choose colors
col <- c("green", "yellow", "red")

# Plot points with x=Longitude and y=Latitude. Colors by contrast category
world_map <- world_map + geom_point(aes(x=df$Long, y=df$Lat, color=df$Categories), size=2) +
  scale_colour_manual(values = col, 
                      name  ="# Contrasts",  
                      breaks = c("Less than 3", "Exactly 3", "More than 3"), 
                      labels = c("Less than 3", "Exactly 3", "More than 3")) +
                      xlab("Longitude") +
                      ylab("Latitude") +
                      labs(title="Demonstratives World Map")

ggsave("world_map.png", plot = world_map, width = 21.6, height = 10.8, units = 'in',  dpi=100)
ggsave("world_map2.png", plot = world_map, width = 28.8, height = 14.4, units = 'in',  dpi=100)

                        
world_map
```



### Attempt at interactive plot
``` {r PlotInter}
library(plotly)


# geo styling
g <- list(
  showland = T,
  landcolor = toRGB("gray90"),
  subunitcolor = toRGB("gray90"),
  countrycolor = toRGB("white")
)

# Make interactive plot!
int_plot <- plot_geo(df, lat = ~Lat, lon = ~Long) %>%
  add_markers(text = ~paste(Lang, paste("GeoDryer: ", GeoDryer), 
                            paste("GeoNichols: ", GeoNichols), 
                            paste("Stock: ", Stock), 
                            paste("Latitude: ", round(Lat, 2)), 
                            paste("Longitude: ", round(Long, 2)), 
                            paste("Nr Contrasts:", Contrasts), 
                            paste("Orientation: ", Orientation), 
                            sep = "<br />"),
              color = ~Categories, colors = col, hoverinfo = "text"
  ) %>%
  add_annotations(text="# Contrasts", xref="paper", yref="paper",
                  x=1.02, xanchor="left",
                  y=0.8, yanchor="bottom",    # Same y as legend below
                  legendtitle=TRUE, showarrow=FALSE ) %>%
  layout(
    title = 'Demonstratives World Map<br />(Number of Contrasts)', 
    xaxis = list(title = 'Longitude'), 
    yaxis = list(title='Latitude'), 
    geo = g,
    legend = list(x = 1.02, y = 0.8, xanchor= 'left', yanchor = "top"),
    autosize = F,
    width = 1600,
    height = 800
  )


# Display plot
# int_plot

# Save plot
htmlwidgets::saveWidget(as_widget(int_plot), "world_map_interactive.html")
```






